### 2022-05-30 paper.Rmd

To thorougly evaluate k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.

The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (c) a simple random allocation of elements to groups. All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step. In the case of k-plus anticlustering, the augmented input data---i.e., the original data as well as the features that were added to incorporate the additional optimization criteria---were subjected to a standardization, which is recommended.

@@

The same local maximum search was applied to optimize each of the five competing anticlustering objectives [Method LCW in @weitz1998]. Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin. That is, for each element, all possible exchanges are simulated and the resulting objective values are stored. After each exchange has been tested, the one exchange is realized that improves the objective the most. No exchange is realized if no improvement is possible. The exchange procedure is repeated for each element. Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements, repeating the exchange procedure. The procedure stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found. To obtain better results, this local maximum search may be initialized multiple times [@spath1986]. The simulation employed five repetitions of the local maximum search.

@@

I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each; this process takes about 3-5 seconds on a contemporary personal computer, depending on how many iterations are needed to find the local maximum. The task was accomplished using the following R code:

@@

Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence. The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**). The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion. Only the argument `variance` is set to `TRUE` by default; the other arguments have to be "turned on". If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.

The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section. The argument `K` describes the number of groups. By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process (this behaviour can be adjusted using the boolean argument `standardize`).

Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features. Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing. It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations. K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations. Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus. The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead maximizes within-group heterogeneity.

@@

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. K-plus anticlustering strives for equal means and variances in all groups, regardless of group size. Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:

```R
anticlustering(
  features,
  K = c(100, 100, 100, 600),
  standardize = TRUE,
  objective = "kplus",
  method = "local-maximum"
)
```

@@

  K <- c(100, 100, 100, 600)
  mean_sd_tabs <- list()
  for (criterion in c("kplus", "variance", "diversity")) {
    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
      features,
      standardize = TRUE,
      K = K,
      objective = criterion,
      method = "local-maximum"
    )
    mean_sd_tabs[[criterion]] <- mean_sd_tab(
      features,
      oasis[[paste(criterion, "group", sep = "_")]]
    )
    mean_sd_tabs[[criterion]] <- cbind(1:length(K), K, mean_sd_tabs[[criterion]])
  }

>>>>

To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.

The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (iii) a simple random allocation of elements to groups. All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step. In the case of k-plus anticlustering, the augmented input data---i.e., the original data as well as the features that were added to incorporate the additional optimization criteria---were subjected to a standardization, which is recommended.

@@

The same local maximum search [Method LCW in @weitz1998] was applied to optimize each of the five competing anticlustering objectives. Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin. That is, for each element, all possible exchanges are simulated and the resulting objective values are stored. After each exchange has been tested, the one exchange is realized that improves the objective the most. No exchange is realized if no improvement is possible. The exchange procedure is repeated for each element. Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements, repeating the exchange procedure. The procedure stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found. To obtain better results, this local maximum search may be initialized multiple times [@spath1986]. The simulation employed five repetitions of the local maximum search.

@@

I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each. The task was accomplished using the following R code:

@@

The code runs in about 10 seconds on a contemporary personal computer and varies depending on how many iterations are needed to find the local maximum. The variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence. The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the variable `features` (**TODO**).

The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion. Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes the variance in addition to the mean. The arguments corresponding to skewness, kurtosis and convariance have to be spefically "turned on". If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector. The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section. The argument `K` describes the number of groups. By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process. This behaviour can be adjusted using the boolean argument `standardize`. The documentation of the `anticlust` package also provides detailed information on how to use the `kplus_anticlustering()` function.

Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features. Table 2 also entails the results that were obtained when applying k-means and diversity anticlustering with the same specifications. It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations. The skewness is also very evenly matched. K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and especially skewness. Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus. The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity. Generally, all methods fulfilled their goals pretty well because the data set they were applied to was rather large.

@@

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. K-plus anticlustering strives for equal means and variances in all groups, regardless of group size. Unequal group sizes can be requested in anticlust by passing the different group sizes to the argument `K`. The following code was used to equalize means and variances between four groups:

```R
kplus_anticlustering(
  features,
  K = c(100, 100, 100, 600),
  method = "local-maximum"
)
```

@@

  K <- c(100, 100, 100, 600)
  mean_sd_tabs <- list()
  for (criterion in c("kplus", "variance", "diversity")) {
    if (criterion == "kplus") {
        oasis[[paste(criterion, "group", sep = "_")]] <-  kplus_anticlustering(
            features,
            K = c(100, 100, 100, 600),
            method = "local-maximum"
        )
    } else {
        oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
        features,
        standardize = TRUE,
        K = K,
        objective = criterion,
        method = "local-maximum"
        )
    }
    mean_sd_tabs[[criterion]] <- mean_sd_tab(
      features,
      oasis[[paste(criterion, "group", sep = "_")]]
    )
    mean_sd_tabs[[criterion]] <- cbind(1:length(K), K, mean_sd_tabs[[criterion]])
  }
